---
title: "COVID-19 Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: ['twitter','facebook','menu']
---

```{r setup, include=FALSE}

#load packages : https://epirhandbook.com/en/new_pages/cleaning.html
if(!require(pacman)) install.packages("pacman", repos = "http://cran.us.r-project.org")

pacman::p_load(tidyverse,
               dplyr, RKaggle,skimr, gtsummary, flextable, dlookr, flexdashboard,
               knitr, DT, ggplot2, openintro, highcharter, rpivotTable, janitor, MASS,
               ggbiplot, lubridate, RColorBrewer, plotly)


#import data
#use strings as factors argument to convert all strings to factors one time
# data <- RKaggle::get_dataset("arashnic/covid19-case-surveillance-public-use-dataset")

data <-  read_csv("COVID-19_Case_Surveillance_Public_Use_Data.csv")

# skim(data)

# str(data)

data <- data |> mutate(epi_year=epiyear(cdc_report_dt), epi_week = epiweek(cdc_report_dt))

overall <- data |> group_by(epi_year,epi_week) |> summarise(cases=n()) 

overall_by_wk <- data |> group_by(epi_year,epi_week, cdc_report_dt) |> summarise(cases=n()) 

hosp_admissions <- data |> filter(hosp_yn == "Yes")
hosp_admissions_count <- nrow(hosp_admissions)
```

# Overview

## Row

### Number of Epi Weeks

```{r}
valueBox(length(overall$epi_week),
         icon="fa-calendar-days")

```

### Number of Cases to Date

```{r}
valueBox(sum(overall$cases),
         icon="fa-hospital")

```

### Number of Admissions to Hospital

```{r}

valueBox(hosp_admissions_count,
         icon="fa-timeline")

```

## Column {data-width="500"}

### COVID-19 Cases in 2020

```{r}

fig <- plot_ly(overall_by_wk, type = "scatter",mode="lines") %>%
  add_trace(x=~cdc_report_dt, y=~cases) %>%
  layout(showlegend = F, title='COVID-19 Cases in 2020',
  xaxis = list(rangeslider = list(visible = T),
               rangeselector=list(
        buttons=list(
        list(count=1, label="1 month", step="month", stepmode="backward"),
        list(count=6, label="6 months", step="month", stepmode="backward"),
        list(count=1, label="YTD", step="year", stepmode="todate"),
        list(count=1, label="1 year", step="year", stepmode="backward"),
        list(step="all")
            )))) %>% 
  layout(
    xaxis = list(zerolinecolor = '#ffff',
                 zerolinewidth = 2,
                 gridcolor = 'ffff'),
    yaxis = list(zerolinecolor = '#ffff',
                 zerolinewidth = 2,
                 gridcolor = 'ffff'),
    plot_bgcolor='#e5efe5', margin = 0.1, width = 600)
fig

```

Column {data-width=500}

### COVID-19 Cases by Age Category

```{r}
col<-brewer.pal (9,"BrBG")
fig.2 <- data |> 
  filter(!is.na(age_group), age_group != "Unknown") |>  #filter out unknown and NA
  group_by(age_group) |>
  summarise(cases=n()) |> arrange(desc(cases)) |>
  hchart('bar',hcaes(x=age_group,y= cases,color=col)) |>  
  hc_add_theme(hc_theme_google()) |> 
    #what users can see when interact with graph
    hc_tooltip(pointFormat = "<b>Number of COVID-19 Cases: <b> {point.y} <b>")  |>
    hc_title(text="Number of COVID-19 Cases per Age Category",
    style=list(fontSize = "20px",fontWeight="bold")) |>
    hc_subtitle(text="Time Period: 2020",
    style=list(fontSize = "10px")) 
fig.2

# 
# data %>% group_by(age_group) |>
#   summarise(cases=n()) %>% arrange(desc(cases))

```

# Data Table (First 1000 Rows)

```{r}

data_subset <- data[1:1000,]

datatable(data_subset,caption="COVID-19 Cases",
          rownames=TRUE,
          filter="top",
          options(list(pageLength=25)))


```

```{=html}
<style>
.rpivotTable {
   overflow: auto;
}
</style>
```

# Pivot Table

### Pivot Table By Age Category

```{r}
rpivotTable(overall_by_wk,
            aggregatorName="cases",
            cols=c("epi_year","epi_week"),
            rows="age_category",
            renderName="Heatmap")
```

# References {data-orientation="columns"}

## Column {data-width = 500}

### References

1.  Covid-19: Case Surveillance + Modelling [Internet]. [cited 2026 May 17]. Available from: <https://kaggle.com/code/afafathar3007/covid-19-case-surveillance-modelling>
2.  Covid-19: Case Surveillance + Modelling [Internet]. [cited 2026 May 17]. Available from: <https://kaggle.com/code/afafathar3007/covid-19-case-surveillance-modelling>
3.  COVID-19 Case Surveillance Public Use Data \| Data \| Centers for Disease Control and Prevention [Internet]. [cited 2026 May 17]. Available from: <https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf/about_data>
4.  Interactive Dashboards in R \| Data Visualization with Popular Plots, Data Tables & Pivot Charts - YouTube [Internet]. [cited 2026 May 17]. Available from: <https://www.youtube.com/watch?v=_a4S4tq62OE&ab_channel=Dr.BharatendraRai>
5.  42 Dashboards with R Markdown – The Epidemiologist R Handbook [Internet]. [cited 2026 May 17]. Available from: <https://epirhandbook.com/en/new_pages/flexdashboard.html>
6.  Find the Perfect Icon for Your Project \| Font Awesome [Internet]. [cited 2026 May 17]. Available from: <https://fontawesome.com/search?q=time>
7.  Time series and date axes in R [Internet]. [cited 2026 May 17]. Available from: <https://plotly.com/r/time-series/>
8.  How to Create Cool Dashboards using R - YouTube [Internet]. [cited 2026 May 17]. Available from: <https://www.youtube.com/watch?v=fkqD9kcvCkU&t=223s&ab_channel=DatawithMiguel>
9.  Create a value box component for a dashboard. — valueBox • flexdashboard [Internet]. [cited 2026 May 17]. Available from: <https://rstudio.github.io/flexdashboard/reference/valueBox.html>
10. Create a value box component for a dashboard. — valueBox [Internet]. [cited 2026 May 17]. Available from: <https://rstudio.github.io/flexdashboard/reference/valueBox.html>
11. Using flexdashboard • flexdashboard [Internet]. [cited 2026 May 17]. Available from: <https://rstudio.github.io/flexdashboard/articles/using.html>
12. Colors in R - Easy Guides - Wiki - STHDA [Internet]. [cited 2026 May 17]. Available from: <https://www.sthda.com/english/wiki/colors-in-r>
13. rOpenSci \| How to Cite R and R Packages [Internet]. [cited 2026 May 17]. Available from: <https://ropensci.org/blog/2021/11/16/how-to-cite-r-and-r-packages/>
14. Gesmann M. Interactive pivot tables with R \| R-bloggers [Internet]. 2015 Mar 31 [cited 2026 May 17]. Available from: <https://www.r-bloggers.com/2015/03/interactive-pivot-tables-with-r/>
15. RyanStochastic. How to set size of rpivotTable in rmarkdown document [Forum post]. Stack Overflow [Internet]. 2021 [cited 2026 May 17]. Available from: <https://stackoverflow.com/q/69008496>
16. Kat. Answer to “How to set size of rpivotTable in rmarkdown document.” Stack Overflow [Internet]. 2022 [cited 2026 May 17]. Available from: <https://stackoverflow.com/a/71975709>
17. danieldavid521. Answer to “How to set size of rpivotTable in rmarkdown document.” Stack Overflow [Internet]. 2021 [cited 2026 May 17]. Available from: <https://stackoverflow.com/a/69016410>

## Column {data-width = 500}

### About Report

Report Generated by: Shardha Ramlal

Date: `r format(Sys.Date(), format= "%B %d, %Y")`

Data Source:

- CDC - <https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf/about_data>

- <https://www.kaggle.com/code/afafathar3007/covid-19-case-surveillance-modelling>

```{r}
#This report was generated on `r format(Sys.Date(), format="%B" %d, %Y")
devtools::session_info()

```
